Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-10-24 thirdPaper/draft/typeConeAssociahedra.tex
coefficients cluster algebras

\newtheorem*{thm:allPolytopalRealizationsCA}{\cref{thm:allPolytopalRealizationsCA}}
\begin{thm:allPolytopalRealizationsCA}
For any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
R_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\principalVariables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \sum_{y \in \principalVariables} \alpha_{x,x',y} \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for all } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$R_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.
\end{thm:allPolytopalRealizationsCA}

Here, $\meshes$ denotes the set of all pairs~$\{x,x'\}$ related by non-initial mesh mutations (\cref{def:meshMutation}), and~$\principalVariables$ the set of cluster variables.
For $\{x,x'\} \in \meshes$ and $y \in \principalVariables$, the coefficient $\alpha_{x,x',y}$ is defined as follows: $\alpha_{x,x',y}$ is set to be~$|b_{xy}|$ if~$y\in X\cap X'$ and $0$ otherwise, where~$X,X'$ are any two clusters such that~$X \ssm \{x\} = X' \ssm \{x'\}$. By the unique exchange relation property, the coefficients~$\alpha_{x,x',y}$ do not depend on the specific choice of clusters~$X,X'$.

@@

For~$\{x,x'\} \in \meshes$ and~$y \in \principalVariables$, we denote by~$\alpha_{x,x',y}$ the coefficient of~$\gvectorFull{\B_\circ}{y}$ in the linear dependence between the $\b{g}$-vectors~$\gvectorFull{\B_\circ}{x}$ and~$\gvectorFull{\B_\circ}{x'}$. In other words, according to \cref{lem:dependenceMesh}, if~${(\B, \cluster)}$ and~${(\B', \cluster')}$ are two adjacent seeds with~$\cluster \ssm \{x\} = \cluster' \ssm \{x'\}$, we have~$\alpha_{x,x',y} = |b_{xy}|$ for~$y \in \cluster \cap \cluster'$ and~$\alpha_{x,x',y} = 0$ otherwise.
The following statement will be shown in \cref{corollary:proof of prop meshMutations}.

@@

\begin{theorem}
\label{thm:allPolytopalRealizationsCA}
For any finite type exchange matrix~$\B_\circ$ (acyclic or not, simply-laced or not), and for any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
R_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\principalVariables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \sum_{y \in \principalVariables} \alpha_{x,x',y} \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for all } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$R_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.
\end{theorem}

\begin{example}
To complement \cref{subsec:typeConeAsso}, let us translate to diagonals and triangulations the result of \cref{thm:allPolytopalRealizationsCA} in type~$A$.
Let~$n \ge 3$.
Consider a convex $(n+3)$-gon whose vertices are labeled modulo~$n+3$ and whose internal and boundary diagonals are denoted by~$\Delta(n) \eqdef \binom{\Z/(n+3)\Z}{2}$.
Consider a triangulation~$T$ of this polygon, formed by the $n+3$ boundary edges and $n$ internal diagonals of the $(n+3)$-gon.
Then for any~$\b{\ell} \in \R_{>0}^{\Delta(n) \ssm T}$, the polytope
\[
R_\b{\ell}(T) \eqdef \Bigset{\b{z} \in \R^{\Delta(n)}}{\begin{array}{l} \b{z} \ge 0 \quad\text{ and }\quad \b{z}_{(a,b)} = 0 \text{ for all } (a,b) \in T \quad\text{ and }\quad \\ \b{z}_{(a,b)} + \b{z}_{(a-1,b-1)} - \b{z}_{(a,b-1)} - \b{z}_{(a-1,b)} = \b{\ell}_{(a,b)} \text{ for all } (a,b) \notin T \end{array}}
\]
is an associahedron whose normal fan is the $\b{g}$-vector fan~$\gvectorFan[T]$ with respect to the triangulation~$T$.
Moreover, the polytopes~$R_\b{\ell}(T)$ for~$\b{\ell} \in \R_{>0}^{\Delta(n) \ssm T}$ describe all polytopal realizations of~$\gvectorFan[T]$.
\end{example}

>>>>

\newtheorem*{thm:allPolytopalRealizationsCA}{\cref{thm:allPolytopalRealizationsCA}}
\begin{thm:allPolytopalRealizationsCA}
For any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
R_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\principalVariables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \sum_{y \in \principalVariables} \coefficient[y][x][x'] \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for all } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$R_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.
\end{thm:allPolytopalRealizationsCA}

Here, $\meshes$ denotes the set of all pairs~$\{x,x'\}$ related by non-initial mesh mutations (\cref{def:meshMutation}), and~$\principalVariables$ the set of cluster variables.
For $\{x,x'\} \in \meshes$ and $y \in \principalVariables$, the coefficient $\coefficient[y][x][x']$ is defined as follows: $\coefficient[y][x][x']$ is set to be~$|b_{xy}|$ if~$y\in X\cap X'$ and $0$ otherwise, where~$X,X'$ are any two clusters such that~$X \ssm \{x\} = X' \ssm \{x'\}$. By the unique exchange relation property, the coefficients~$\coefficient[y][x][x']$ do not depend on the specific choice of clusters~$X,X'$.

@@

For~$\{x,x'\} \in \meshes$ and~$y \in \principalVariables$, we denote by~$\coefficient[y][x][x']$ the coefficient of~$\gvectorFull{\B_\circ}{y}$ in the linear dependence between the $\b{g}$-vectors~$\gvectorFull{\B_\circ}{x}$ and~$\gvectorFull{\B_\circ}{x'}$. In other words, according to \cref{lem:dependenceMesh}, if~${(\B, \cluster)}$ and~${(\B', \cluster')}$ are two adjacent seeds with~$\cluster \ssm \{x\} = \cluster' \ssm \{x'\}$, we have~$\coefficient[y][x][x'] = |b_{xy}|$ for~$y \in \cluster \cap \cluster'$ and~$\coefficient[y][x][x'] = 0$ otherwise.
The following statement will be shown in \cref{corollary:proof of prop meshMutations}.

@@

\begin{theorem}
\label{thm:allPolytopalRealizationsCA}
For any finite type exchange matrix~$\B_\circ$ (acyclic or not, simply-laced or not), and for any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
R_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\principalVariables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \sum_{y \in \principalVariables} \coefficient[y][x][x'] \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for all } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$R_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.
\end{theorem}

\begin{example}
To complement \cref{subsec:typeConeAsso}, let us translate to diagonals and triangulations the result of \cref{thm:allPolytopalRealizationsCA} in type~$A$.
Let~$n \ge 3$.
Consider a convex $(n+3)$-gon whose vertices are labeled modulo~$n+3$ and whose internal and boundary diagonals are denoted by~$\Delta(n) \eqdef \binom{\Z/(n+3)\Z}{2}$.
Consider a triangulation~$T$ of this polygon, formed by the $n+3$ boundary edges and $n$ internal diagonals of the $(n+3)$-gon.
Then for any~$\b{\ell} \in \R_{>0}^{\Delta(n) \ssm T}$, the polytope
\[
R_\b{\ell}(T) \eqdef \Bigset{\b{z} \in \R^{\Delta(n)}}{\begin{array}{l} \b{z} \ge 0 \quad\text{ and }\quad \b{z}_{(a,a+1)} = 0 \text{ for all } a \in \Z/(n+3)\Z \quad\text{ and }\quad \\ \b{z}_{(a,b)} + \b{z}_{(a-1,b-1)} - \b{z}_{(a,b-1)} - \b{z}_{(a-1,b)} = \b{\ell}_{(a,b)} \text{ for all } (a,b) \notin T \end{array}}
\]
is an associahedron whose normal fan is the $\b{g}$-vector fan~$\gvectorFan[T]$ with respect to the triangulation~$T$.
Moreover, the polytopes~$R_\b{\ell}(T)$ for~$\b{\ell} \in \R_{>0}^{\Delta(n) \ssm T}$ describe all polytopal realizations of~$\gvectorFan[T]$.
\end{example}
